Using cosmic variance to constrain the dark matter halo mass of Lyman-alpha emitting galaxies at $z$=3.1

### 2013-05-27 paper/ms.tex
updated observational constraints section

\section{Methodology}
In this \documentname we constrain the typical mass of dark matter halos
hosting LAES at $z=3.1$. Our model is based on the number density
information obtained in the recent large scale survey presented by XXX
where XXX LAEs are detected over 7 fields of $\sim 46 \times
35$Mpc$^{2}h^{-2}$ in area comoving in area corresponding to observed
fields of $XXX$ deg$^{2}$.

Both the spatial distribution and luminosities of the galaxies have,
at least in a statistical sense, informatio to constrain theoretical
model of LAEs. The most detailed theoretical models are also faced to
diverse physical and astrophysical uncertainties in obtaining
statistical prediction for the Lyman-alpha line. This uncertainties
are largest impediment to consruct an ab-iitio model for LAEs.

In this \documentname we want to step back and reduce the complexity of our
model, with the sole objective of reproducing the cosmic variance in
the number density of LAEs. Afterwards we will interpret the
implications of this result for physical models for Lyman-alpha
emitting galaxies.

Our model is based on the predictions of a large volume high
resolution N-body simulation describing the gravitational dynamics of
dark matter. We do not have an strong bias towards the theoretical
expectation of what the mass of the dark matter halo hosting the
galaxy should be.  Instead, we fully explore the parameter space of
our simplified model. The only cut we impose is that observed LAEs do
not reside in dark matter halos with masses less than 10$^{10}$
\hMsun [citation].

In the following subsections we describe the most relevant features of
the observational data, the N-body simulation we use, our model and
its parameters together with the method to compare its predictions
against observations.

@@

Our observational reference are the recently published results of a
panoramic survey of LAES at $z=3.1$ by \cite{Yamada2012}. This survey
was conducted with the Subaru 8.2m telescope and the Subaru Prime
Focus Camera, which has a field of view covering $34\times 27$ arcmin,
corresponding to a comoving scale of $46\times35$ Mpc $h^{-1}$ at
$z=3.09$. The narrow band filter is centered at $4977$ \AA with a
$77$\AA width, corresponding to the redshift range $z=3.062-3.125$ and
$41$ Mpc $h^{-1}$ comoving scale for the detection of the
Lyman-$\alpha$ line centered at $z=3.09$.

The choice to have only one the data from Yamada et al 2012 as
reference was  made because their surveys is the largest in area with
a set of homogenous conditions that define the LAE sample. Other
surveys by XXX an XXX that cover similar regions, but they use
different criteria on the equivalent width (EW) cuts to construct the
LAE samples. Different cuts in the EW can change the number of LAEs to
be included in the catalog. This cuts have an impact on the fainter
LAEs which are more abundante than brighter ones. Different
definitions of the EW cuts can yield number densities different by a
factor of two [REF, I think Yamada has some numbers].

The survey covered four independent fields. The first is the SSA22
field of $1.38$ deg$^2$ with $1394$ detected LAEs, this field has been
known to harbor a region with a large density excess of galaxies. The
second observed region is composed by the fields Subaru/{\it
  XMM-Newton} Deep Survey (SXDS)-North, -Center and -South, with a 
total of $0.58$ deg$^2$ and $386$ LAEs. The third and fourth fields
are the Subaru Deep Field (SDF) with $0.22$ deg$^2$ and $196$ LAEs,
and the fild arotund the Great Observatory Optical Deep Survey North
(GOODS-N) with $0.24$ deg$^2$ and $185$ LAEs. In Table 1 we summarize
the values we use in throughout this paper for the each field, covered
area, measured surface LAE number density and inferred number volume
density.

@@

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/match_large_correlation_selected_models.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/match_full_correlation_selected_models.pdf}
\end{center} f
\caption{ mean ACFs   and their correponding standard deviation (error
  bars)  of some selected models in different mass ranges over the 7
  subfields  f the SSA22 field (left) and the entire 12 field sample
  (rigth) using thematch
  configuration. \label{figure:correlation_match} }  
  \end{figure*}

\begin{figure*}
\begin{center}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/random_large_correlation_selected_models.pdf}
\hspace{5mm}
\includegraphics[width=0.46\linewidth,angle=0]{./plots/random_full_correlation_selected_models.pdf}
\end{center} 
\caption{ mean ACFs   and their correponding standard deviation (error
  bars)  of some selected models in different mass ranges over the 7
  subfields of the SSA22 field (left) and the entire 12 field sample
  (rigth) using the  random
  configuration. \label{figure:correlation_random} }

\end{figure*}

>>>>

\section{Methodology}

Our method to constrain the typical mass of a dark matter halos
hosting LAEs at $z=3.1$ is based on the comparison of observational
results on the surface number density and the predictions of a simple
model that uses the outputs from cosmological N-body simulations.

In the next subsections we describe in detail the four key elements of
this workflow. First, we present the observations we take as a
benchmark. Second, the N-body simulation and the halo catalogs we
use. Third, the simplified model that allows us to translated halo
catalogs into mock LAE observations. Fourth, the statistics we use to
compare observational results against our theoretical predictions.

%In this \documentname we constrain the typical mass of dark matter halos
%hosting LAES at $z=3.1$. Our model is based on the number density
%information obtained in the recent large scale survey presented by XXX
%where XXX LAEs are detected over 7 fields of $\sim 46 \times
%35$Mpc$^{2}h^{-2}$ in area comoving in area corresponding to observed
%fields of $XXX$ deg$^{2}$.

%Both the spatial distribution and luminosities of the galaxies have,
%at least in a statistical sense, informatio to constrain theoretical
%model of LAEs. The most detailed theoretical models are also faced to
%diverse physical and astrophysical uncertainties in obtaining
%statistical prediction for the Lyman-alpha line. This uncertainties
%are largest impediment to consruct an ab-iitio model for LAEs.

%In this \documentname we want to step back and reduce the complexity of our
%model, with the sole objective of reproducing the cosmic variance in
%the number density of LAEs. Afterwards we will interpret the
%implications of this result for physical models for Lyman-alpha
%emitting galaxies.

%Our model is based on the predictions of a large volume high
%resolution N-body simulation describing the gravitational dynamics of
%dark matter. We do not have an strong bias towards the theoretical
%expectation of what the mass of the dark matter halo hosting the
%galaxy should be.  Instead, we fully explore the parameter space of
%our simplified model. The only cut we impose is that observed LAEs do
%not reside in dark matter halos with masses less than 10$^{10}$\hMsun
%[citation].

%In the following subsections we describe the most relevant features of
%the observational data, the N-body simulation we use, our model and
%its parameters together with the method to compare its predictions
%against observations.

@@

The observational benchamark we use in this paper is the LAE number
density information at $z=3.1$ obtained by the panoramic narrow-band
survey presented by \cite{Yamada2012} from a survey 
%Our observational reference are the recently published results of a
%panoramic survey of LAES at $z=3.1$ by \cite{Yamada2012}. 
%This survey
conducted with the Subaru 8.2m telescope and the Subaru Prime
Focus Camera, which has a field of view covering $34\times 27$ arcmin,
corresponding to a comoving scale of $46\times35$ Mpc $h^{-1}$ at
$z=3.09$. The narrow band filter is centered at $4977$ \AA with a
$77$\AA width, corresponding to the redshift range $z=3.062-3.125$ and
$41$ Mpc $h^{-1}$ comoving scale for the detection of the
Lyman-$\alpha$ line centered at $z=3.09$. The authors report a total of
$2161$ LAEs with an observed equivalent width larger than $190$\AA
over a total survey area of $2.42$ deg$^{2}$, this corresponds to an
averaga surface number density of $0.20\pm 0.01$ arcmin$^{-2}$.

The survey covered four independent fields. The first is the SSA22
field of $1.38$ deg$^2$ with $1394$ detected LAEs, this field has been
known to harbor a region with a large density excess of galaxies. The
second observed region is composed by the fields Subaru/{\it
  XMM-Newton} Deep Survey (SXDS)-North, -Center and -South, with a 
total of $0.58$ deg$^2$ and $386$ LAEs. The third and fourth fields
are the Subaru Deep Field (SDF) with $0.22$ deg$^2$ and $196$ LAEs,
and the fild arotund the Great Observatory Optical Deep Survey North
(GOODS-N) with $0.24$ deg$^2$ and $185$ LAEs. 
%In Table 1 we summarize
%the values we use in throughout this paper for the each field, covered
%area, measured surface LAE number density and inferred number volume
%density.

There is abundant observational work done on LAEs at redshift $z=3.1$
\citep{Kudritzki2000,Matsuda2005,Gawiser2007,Nilsson2007,Ouchi2008}.
However, we decide to focus on the data from \cite{Yamada2012} because
it has the largest covered area with homogeneous intrumentational
conditions (telescope, narrow band filter), data reduction pipeline
and conditions to construct the LAE catalog. This ensures that the
number density variations among fields are due only to
astrophysical reasons and not different observational conditions or
criteria to construct the catalogs.

%For instance, having
%different selection criteria on the equivalent width (EW) can change
%the number of LAEs included in the catalog, yielding number densities
%different by a factor of two   
%different criteria on the equivalent width (EW) cuts to construct the
%LAE samples. Different cuts in the EW can change the number of LAEs to
%be included in the catalog. This cuts have an impact on the fainter
%LAEs which are more abundante than brighter ones. Different
%definitions of the EW cuts can yield number densities different by a
%factor of two [REF, I think Yamada has some numbers].

%Other
%surveys by XXX an XXX that cover similar regions, but they use
%different criteria on the equivalent width (EW) cuts to construct the
%LAE samples. Different cuts in the EW can change the number of LAEs to
%be included in the catalog. This cuts have an impact on the fainter
%LAEs which are more abundante than brighter ones. Different
%definitions of the EW cuts can yield number densities different by a
%factor of two [REF, I think Yamada has some numbers].

@@

%\begin{figure*}
%\begin{center}
%\includegraphics[width=0.46\linewidth,angle=0]{./plots/match_large_correlation_%selected_models.pdf}
%\hspace{5mm}
%\includegraphics[width=0.46\linewidth,angle=0]{./plots/match_full_correlation_s%elected_models.pdf}
%\end{center} f
%\caption{ mean ACFs   and their correponding standard deviation (error
%  bars)  of some selected models in different mass ranges over the 7
%  subfields  f the SSA22 field (left) and the entire 12 field sample
%  (rigth) using thematch
%  configuration. \label{figure:correlation_match} }  
%  \end{figure*}

%\begin{figure*}
%\begin{center}
%\includegraphics[width=0.46\linewidth,angle=0]{./plots/random_large_correlation_selected_models.pdf}
%\hspace{5mm}
%\includegraphics[width=0.46\linewidth,angle=0]{./plots/random_full_correlation_selected_models.pdf}
%\end{center} 
%\caption{ mean ACFs   and their correponding standard deviation (error
%  bars)  of some selected models in different mass ranges over the 7
%  subfields of the SSA22 field (left) and the entire 12 field sample
%  (rigth) using the  random
%  configuration. \label{figure:correlation_random} }   
%\end{figure*}
